Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-23 robdml.tex
tex

%%%%%%%%% ABSTRACT
\begin{abstract}
	% background
	Owing to the security implications of adversarial vulnerability, the
	adversarial robustness of deep metric learning models have to be improved
	via defense.
	% insight
	In order to avoid model collapse due to excessively hard examples, the
	existing methods refrain from using the min-max adversarial training, but
	instead learns from a weak adversary inefficiently.
	% our starting point
	Conversely, we argue that a model can learn from a strong adversary without
	collapse, where the key is triplet hardness.
	% HM framework
	To this end, Hardness Manipulation is proposed to adversarially perturb the
	training triplet to increase its hardness into a specified level, by
	referencing another benign triplet or a pseudo-hardness function.
	% GA Adversary
	Based on this, Gradual Adversary is proposed as a family of pseudo-hardness
	functions to overcome problems of referencing other benign triplets, and
	achieve a better balance in the learning objective.
	% Intra-Class Structure
	Besides, a penalty on Intra-Class Structure among benign and adversarial
	examples can further improve model robustness, which is neglected by
	previous methods.
	% Experiment and Conclusion
	Comprehensive experimental results on commonly used deep metric
	learning datasets suggest that the proposed method, although simple in
	form, overwhelmingly outperforms the state-of-the-art defense methods in
	terms of training time cost, robustness, as well as recall performance on
	benign examples.
%
\end{abstract}

@@

% existing methods & problem
Existing defense methods for DML~\cite{advrank,robrank} are adversarial
training-based, inspired by Madry's \emph{min-max} adversarial
training~\cite{madry} because it is one of the most effective methods for deep
neural network classifiers.
%
Specifically, Madry's method involves a inner problem to \emph{maximize} the
loss by perturbing the inputs into adversarial examples, and an outer problem
to \emph{minimize} the loss by updating the model parameters.
%
However, in order to avoid model collapse, the existing DML defenses refrain
from directly adopting such paradigm, but instead use some indirect and
relatively weak objectives to merely increase the loss value to a certain
level.
%
As a result, the model is not learning from a strong adversary, and suffers
from low efficiency in gaining robustness within limited time. 
%
Furthermore, adversarial training already suffers from significant training
time cost (compared to the normal training process).
%
Hence, time efficiency in gaining adversarial robustness
is an inevitable major issue.

% we argue

@@

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified destination hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
destination hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified destination hardness level.
%
But, how hard should such ``destination hardness'' be?

% how hard? benign destination reference

Recall that the model is already prone to collapse with excessively hard
benign triplets, let alone adversarial examples.
%
Thus, intuitively, the destination hardness can be that of another benign
triplet which is harder than the given triplet (\eg, a Semihard~\cite{facenet}
triplet).

@@

\begin{align}
	&
	\sign \big\{
		-\frac{\partial}{\partial r} \big\| \min(0, 
		\tilde{H}_\mathsf{S} - H_\mathsf{D}) \big\|_2^2
	\big\}
	\\
	=&
	\sign \big\{
		-2(\tilde{H}_\mathsf{S}-H_\mathsf{D})
		\frac{\partial }{\partial r} \tilde{H}_\mathsf{S} \big\}
	=
	\sign \big\{
		\frac{\partial}{\partial r} \tilde{H}_\mathsf{S}
		\big\}
\end{align}

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
	% background
	Owing to the security implications of adversarial vulnerability, the
	adversarial robustness of deep metric learning models have to be improved
	via defense.
	% insight
	In order to avoid model collapse due to excessively hard examples, the
	existing methods refrain from using the min-max adversarial training, but
	instead learns from a weak adversary inefficiently.
	% our starting point
	Conversely, we argue that a model can learn from a strong adversary without
	collapse, where the key is triplet hardness.
	% HM framework
	To this end, Hardness Manipulation is proposed to adversarially perturb the
	training triplet to increase its hardness into a specified level, according
	to another benign triplet or a pseudo-hardness function.
	% GA Adversary
	Based on this, Gradual Adversary is proposed as a family of pseudo-hardness
	functions to overcome problems of referencing other benign triplets, and
	achieve a better balance in the learning objective.
	% Intra-Class Structure
	Besides, a penalty on Intra-Class Structure among benign and adversarial
	examples can further improve model robustness, which is neglected by
	previous methods.
	% Experiment and Conclusion
	Comprehensive experimental results on commonly used deep metric
	learning datasets suggest that the proposed method, although simple in
	form, overwhelmingly outperforms the state-of-the-art defense methods in
	terms of training time cost, robustness, as well as recall performance on
	benign examples.
%
\end{abstract}

@@

% existing methods & problem

Existing defense methods for DML~\cite{advrank,robrank} are adversarial
training-based, inspired by Madry's \emph{min-max} adversarial
training~\cite{madry} because it is one of the most effective methods for deep
neural network classifiers.
%
Specifically, Madry's method involves a inner problem to \emph{maximize} the
loss by perturbing the inputs into adversarial examples, and an outer problem
to \emph{minimize} the loss by updating the model parameters.
%
However, in order to avoid model collapse, the existing DML defenses refrain
from directly adopting such paradigm, but instead use some indirect and
relatively weak objectives to merely increase the loss value to a certain
level.
%
As a result, the model is not learning from a strong adversary, and suffers
from low efficiency in gaining robustness within limited time. 
%
Furthermore, adversarial training already suffers from significant training
time cost (compared to the normal training process).
%
Hence, time efficiency in gaining adversarial robustness
is an inevitable major issue.

% we argue / insight

@@

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified \emph{destination}
hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
\emph{destination} hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified \emph{destination} hardness level.
%
But, how hard should such ``\emph{destination} hardness'' be?

% how hard? benign destination

Recall that the model is already prone to collapse with excessively hard benign
triplets, let alone adversarial examples.
%
Thus, intuitively, the \emph{destination} hardness can be that of another
benign triplet which is harder than the given triplet (\eg, a
Semihard~\cite{facenet} triplet).
%
However, in the late phase of training, the expectation of the loss, as well as
the corresponding \emph{hardness} of a given triplet will be small, which means
the \emph{destination} hardness will be close to that of the sample triplet for
training.
%
This leads to weak attacks, and inefficient learning from a weak adversary.
%
Besides, we speculate a strong adversary in the early phase of DML training may
hinder the model from learning good embeddings, and hence influence the
eventual performance on benign examples.
%
Based on these, a \emph{destination} hardness calculated from benign examples
may not be the best choice.
%
A better \emph{destination} hardness should be able to balance the training
objectives in the early and late phases of training.

@@

\begin{align}
	&
	\sign\big\{
		-\frac{\partial}{\partial r} \big\| \min(0, 
		\tilde{H}_\mathsf{S} - H_\mathsf{D}) \big\|_2^2
	\big\}
	\\
	=&
	\sign\big\{
		-2(\tilde{H}_\mathsf{S}-H_\mathsf{D})
		\frac{\partial }{\partial r} \tilde{H}_\mathsf{S} \big\}
	=
	\sign\big\{
		\frac{\partial}{\partial r} \tilde{H}_\mathsf{S}
		\big\}
\end{align}
